---
title: "Normalization using Seurat"

output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes

knit: (function(input_file, encoding) {
  out_dir <- '../html';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, '01_01_data_normalization_seurat.html'))})
---


#### Define Libraries
```{r echo=TRUE, message=FALSE, warning=FALSE, tidy=TRUE}
library("stringr")
library("ggplot2")
library("Seurat")
library("tidyverse")
library("RColorBrewer")
```

#### Define Global Path
```{r tidy=TRUE}
dir.wrk <- str_replace(getwd(), "/scripts", "")
dir.data <- file.path(dir.wrk, "data/scRNAseq-afterCellRanger/filtered")
dir.des <- file.path(dir.wrk, "data/design")
dir.output <- file.path(dir.wrk, "data/processed_data")
```


#### Get Count Matrix
```{r}
#dirs.batch <- list.dirs(dir.data, full.names = TRUE, recursive = FALSE)

# FUNCTION: myRead10X() ---
myRead10X <- function(dir.data, sampleid) {
  Read10X(data.dir = file.path(dir.data, sampleid))
}

# Load Sparse Matrix ---
sp.mat <- myRead10X(dir.data, sampleid="0661-2_CD138Plus")

# Initialize the Seurat object with the raw (non-normalized data) ---
mm.obj <- CreateSeuratObject(counts = sp.mat, 
                             min.cells = 3,
                             min.features = 100, 
                             project = "0661-2_CD138Plus")
```


#### Get mitochondrial genes stats
```{r}
mm.mito.genes <- grep(pattern = "^MT-", x = rownames(x = mm.obj@assays$RNA@data), value = TRUE)
mm.percent.mito <- Matrix::colSums(mm.obj@assays$RNA@data[mm.mito.genes, ]) / Matrix::colSums(mm.obj@assays$RNA@data)

mm.obj <- AddMetaData(object = mm.obj, metadata = mm.percent.mito,col.name = "percent.mito")

grep("^MT-", rownames(mm.obj@assays$RNA@data), value = TRUE)
```


#### Get QC metrics
```{r}
plot0 <- VlnPlot(object = mm.obj, features= c("nFeature_RNA", "nCount_RNA", "percent.mito"), cols=3)
plot0

plot1 <- FeatureScatter(mm.obj, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(mm.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
#CombinePlots(plots = list(plot1, plot2))
plot1
plot2
```



#### Filter based on QC metrics
```{r}
mm.obj <- subset(mm.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 3500 & percent.mito < 0.1)

plot0 <- VlnPlot(object = mm.obj, features= c("nFeature_RNA", "nCount_RNA", "percent.mito"), cols=3)
plot0

plot1 <- FeatureScatter(mm.obj, feature1 = "nCount_RNA", feature2 = "percent.mito")
plot2 <- FeatureScatter(mm.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
#CombinePlots(plots = list(plot1, plot2))
plot1
plot2
```


#### Normalizing the data
```{r}
mm.obj <- NormalizeData(mm.obj, normalization.method = "LogNormalize", scale.factor = 10000)

expr <- mm.obj[["RNA"]]@data

# Normalized Expression Data can be found in 
#mm.obj[["RNA"]]@data[1:10,1:10]
```

#### WRITE OUTPUT 
```{r}
#file.expr <- file.path(dir.output, "normalized_expr_0661-2_CD138Plus.tsv")
#write.table(expr, file.expr, sep="\t", row.names=TRUE, col.names=NA, quote=FALSE)
```


#### Identification of highly variable features (feature selection)
```{r}
mm.obj <- FindVariableFeatures(mm.obj, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(mm.obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(mm.obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)

plot1
plot2
#CombinePlots(plots = list(plot1, plot2))
```


#### Scale Data 
```{r}
all.genes <- rownames(mm.obj)
mm.obj <- ScaleData(mm.obj, features = all.genes)

expr.scale <- mm.obj[["RNA"]]@scale.data
```

#### Perform linear dimensional reduction
```{r}
mm.obj <- RunPCA(mm.obj, features = VariableFeatures(object = mm.obj))
```

#### Visualize PCA Features
```{r}
VizDimLoadings(mm.obj, dims = 1:2, reduction = "pca")
DimPlot(mm.obj, reduction = "pca")
```

#### PCA Feature Heatmaps
```{r fig.width=8, fig.height=8}
DimHeatmap(mm.obj, dims = 1:15, cells = 500, balanced = TRUE)
```

#### Determine 'dimentionality' of the dataset
```{r}
mm.obj <- JackStraw(mm.obj, num.replicate = 100)
mm.obj <- ScoreJackStraw(mm.obj, dims = 1:20)

### JackStraw PLOT
JackStrawPlot(mm.obj, dims = 1:20)

### Elbow Plot
ElbowPlot(mm.obj)
```



#### Cluster cells
```{r}
mm.obj <- FindNeighbors(mm.obj, dims = 1:10)
mm.obj <- FindClusters(mm.obj, resolution = 0.5)

# Look at cluster IDs of the first 5 cells
head(Idents(mm.obj), 5)
```


#### Run Non-dimentional Reduction (tSNE)
```{r}
mm.obj <- RunTSNE(object = mm.obj, dims.use = 1:15)
```

#### PLOT t-SNE
```{r fig.width=5, fig.height=5}
# COLOR PALETTE ---
jColFun <- colorRampPalette(brewer.pal(n = 9, "Set1"))

# note that you can set do.label=T to help label individual clusters
#TSNEPlot(object = hgg.obj, pt.size = 3, cols=jColFun(8))
DimPlot(object = mm.obj, pt.size = 3, cols=jColFun(12))
```
